---
title: "Automatic Metabolon Analysis"
output: html_notebook
---

# General objective

## Data input and output
Inputs : 
1) Metabolomics raw data is submitted (extracted from metabolon)
2) Patients informations (extracted from metabolon)

Ouputs :  
1)  PCA (condition, age, sex, BMI, HIV status)
2)  PLS-DA (supervized pca)
3)  Two Sample T-Test (Welch's T-test) HIV / HIV + met
4)  Biosigner results
5)  Feature selection using random forest
6)  ANCOVA/ linear regression

# Outline:
##  1) Parsing input file
##  2) Select metabolites with HMDB (filtration)
##  3) PCA
##  4) PLS-DA
##  5) Welch's T-test
##  6) Biosigner
##  7) Feature selection
##  8) ANCOVA/ linear regression

### clean environment
```{r}
rm(list=ls())
```


### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/India_Cameroon/")) 
```

# loading libraries
# loading libraries
```{r, message=FALSE}
library(RColorBrewer)
library(ggfortify)
library(ropls)
library(factoextra)
library(dplyr)
library(biosigner)
library(caret)
library(randomForest)
library(mlbench)
library(RColorBrewer)
library(gplots)
library(reshape2)
library(xlsx)
library(UpSetR)
library(pROC)
```

```{r, message=FALSE}
source("src/India/ML.R")
source("src/India/usefull_functions.R")
```

## Study individual info
```{r}
name_ana <- "India"
data_2 <- read.delim("processing/India/India_clean_data.txt", row.names = 1)
info <- read.xlsx("data/India/India_filter.xlsx",2)

condition <- info$condition

table_condition <-cbind(names(data_2), as.vector(condition))

n_c1 <- 22
n_c2 <- 22
```

##  1) Statistical tests

```{r}
results_man_withney <- my_man_whithney(data_2, 22, 22, "HC", "PLHIV")
```

## biosigner
```{r}
x <- data_2
y <- as.factor(c(rep("HC", 22), rep("PLHIV", 22)))
path_fig <- paste0("results/ML/results_biosigner_hd4.pdf")
#pdf(path_fig)
#diaSign <- biosign(x, y, bootI = 50)
#dev.off()
```

## PLS-DA : feature selection

# PLS-DA

```{r}
# apply it
#x_scaled <- scale(x)

#2. unit-variance scalling

path_fig <- paste0("results/", name_ana, "/PLSDA/", name_ana, 
                   "_PLS-DA_HC_PLHIV.pdf")
pdf(path_fig)
met.plsda <- opls(t(x), y)
dev.off()

vipVn <- data.frame(getVipVn(met.plsda))
vipVn$variable <- rownames(vipVn)
```

```{r}
#ropls.plot(met.plsda, plottype = "score", xvar = "p1", yvar = "p2", hotelling = FALSE,
#  ellipse = TRUE, col.var = NULL, col.pca = NULL)
#ggsave("test_plsda.pdf")
```

# x scores
```{r}
path_fig <- paste0("results/", name_ana, "/PLSDA/", name_ana, 
                   "_PLS-DA_HC_PLHIV_PCA.pdf")
pdf(path_fig)
plot(met.plsda, typeVc = "x-score", parLabVc = rep(".",44), parCexN = 3)
dev.off()
```

```{r}
vipVn <- arrange(vipVn, desc(getVipVn.met.plsda.))
names(vipVn)

#vipVn <- vipVn[1:50,]
```

```{r}
pdf(paste0("results/", name_ana, "/figures/Vip_PLS-DA.pdf"), width = 30)
p<-ggplot(data=vipVn, aes(x=reorder(variable, -getVipVn.met.plsda.), y=getVipVn.met.plsda.)) + theme_classic()+
  geom_bar(stat="identity", fill="#d5e1df", size = 0.01)+ theme(text = element_text(size=4), axis.text.x = element_text(angle = 75, hjust = 1))+ labs(y="VipVn", x = "Metabolites")+ geom_hline(yintercept=1, linetype="dashed", color = "red")
p
dev.off()
```

## select top 30 features
```{r}
vipVn <- vipVn[1:30,]
```

```{r}
features_PLS_DA <- vipVn$variable
```

## 2) random forest with cross validation
```{r}
table_cond <- data.frame(Group = y, t(x))
```

1) Run normal random forest
```{r}
set.seed(123)
fit <- randomForest(Group ~ ., data = table_cond, na.action = na.roughfix)
fit
```

## feature selection using Boruta

```{r}
table_cond <- data.frame(Condition = y, t(x))
```

```{r}
data <- feature_select_Boruta(table_cond, name_ana)
```

```{r}
names(data)[ncol(data)] <- "Group"
```

2) Cross validation
# cross-validation
# mtry tunning
# number of trees tunning

```{r}
# tune mtry parameter : check results for value of mtry between 1 and 15
set.seed(1)
tunegrid <- expand.grid(.mtry = (1:15)) 

# 10 -Fold cross validation
control <- trainControl(method='repeatedcv', 
                        number=10, 
                        repeats=3,
                        savePredictions = TRUE,
                        ## Estimate class probabilities
                        classProbs = TRUE,
                        ## Evaluate performance using 
                        ## the following function
                        summaryFunction = twoClassSummary)


set.seed(123)
#Number randomely variable selected is mtry

rf_default <- train(as.factor(Group)~., 
                    data=data, 
                    method='rf', 
                    metric = 'ROC',
                    tuneGrid=tunegrid, 
                    trControl=control, importance = TRUE, ntree = 1000,
                    )

# keep final model
rf <- rf_default$finalModel

rf
# save final model
file_name <- paste0("results/India/ML/model_rf_2_", name_ana)
saveRDS(rf, file_name)
```

```{r}
rf <- readRDS("results/India/ML/model_rf_2_India")
```

```{r}
library(randomForest)
```

```{r}
test <- results_man_withney
```

```{r}
names(test)[1] <- "variable"
```

```{r}
var_importance <- randomForest::importance(rf)
names <- rownames(var_importance)
var_importance <- data.frame(MeanDecreaseAccuracy = var_importance[,4], variable = names)
var_importance <- arrange(var_importance, MeanDecreaseAccuracy)

position <- var_importance$variable

var_importance <- merge(var_importance, test, by = "variable")

significance <- ifelse(var_importance$Log2FC > 0, "up-regulated", "down-regulated")
significance <- as.factor(significance)
  
var_importance$Log2FC <- NULL
  
var_importance <- arrange(var_importance, MeanDecreaseAccuracy)

var_importance$Significance <- significance
  
path_fig <- paste0("results/", name_ana, "/figures/", name_ana,
                   "_top_variable_importance.pdf")

theme_set(theme_classic())
ggplot(var_importance, aes(x=variable, y=MeanDecreaseAccuracy, color = Significance)) + geom_point(aes(colour=significance, fill=significance), size=3, shape = 21, color = "black")+ 
  scale_fill_manual(values=c('#0066cc','#ff0000'))+ geom_segment(aes(x=variable, 
                     xend=variable, 
                     y=min(MeanDecreaseAccuracy), 
                     yend=max(MeanDecreaseAccuracy)), 
                 linetype="dashed", 
                 size=0.1, color="black")+ scale_x_discrete(limits = position) +   # Draw dashed lines
  labs(title="Variable importance") +  
  coord_flip()
ggsave(path_fig, width = 10)
```

```{r}
selectedIndices <- rf_default$pred$mtry == 1
```


```{r}
g <- ggplot(rf_default$pred[selectedIndices, ], aes(m=HC, d=factor(obs, levels = c("HC", "PLHIV")))) + geom_roc(n.cuts=0) + coord_equal() + style_roc(theme = theme_grey)
g <- g + annotate("text", x = .75, y = .25, label = paste("AUC =", round(calc_auc(g)$AUC, 2)))

ggsave("results/India/figures/India_ROC_hd4_10_cross_validate.pdf")
```
## confusion matrix

```{r}
table <- rf_default$pred[selectedIndices, ][,1:2]
```

```{r}
table <- as.data.frame(table(table))
```
```{r}
ggplotConfusionMatrix <- function(m){
  p <-
    ggplot(data = m ,
           aes(x = obs, y = pred)) +
    geom_tile(aes(fill = log(Freq)), colour = "white") +
    scale_fill_gradient(low = "white", high = "steelblue", limits=c(0, 10)) +
    geom_text(aes(x = obs, y = pred, label = Freq)) +
    theme(legend.position = "none") 
  ggsave("results/India/figures/India_confusion_matrix_india.pdf")
  return(p)
}
```

```{r}
ggplotConfusionMatrix(table)
```
```{r}
list_cross_val <- var_importance$variable
```

# upset plot

```{r}
library(UpSetR)
```

```{r}
results_man_withney <- results_man_withney[results_man_withney$FDR < 0.1,]
```

```{r}
listInput <- list(as.vector(list_cross_val),
                   features_PLS_DA,
                   results_man_withney$Accession)
names(listInput) <- c("RandomForest", "PLS-DA", "Man Withney Test")
```

```{r}
pdf(file="results/India/figures/upset_plot_India.pdf", width=11, height=6) # or other device
upset(fromList(listInput), order.by = "freq",
      main.bar.color = "#000033",
      sets.bar.color=c("maroon","#3939ac","#00b33c"),
      text.scale = 2, point.size
 = 4)
dev.off()
```

```{r}
path_table <- paste0("results/", name_ana, "/", 
                     name_ana, "_metabolites_randomforest.txt")

list_cross_val <- read.delim(path_table)
```

```{r}
path_table <- paste0("results/", name_ana, "/", 
                     name_ana, "_PLS-DA.txt")

features_PLS_DA <- read.delim(path_table)
```

```{r}
path_table <- paste0("results/", name_ana, "/", 
                     name_ana, "_metabolites_welshTtwotest.txt")
features_man_withney <- read.delim(path_table)
```

```{r}
listInput <- list(as.vector(list_cross_val$x),
                   as.vector(features_PLS_DA$x),
                   as.vector(features_man_withney$x))
names(listInput) <- c("RandomForest", "PLS-DA", "Man Withney Test")
```

```{r}
pdf(file="results/India/figures/upset_plot_India_2.pdf", width=11, height=6) # or other device
upset(fromList(listInput), order.by = "freq",
      main.bar.color = "#000033",
      sets.bar.color=c("maroon","#3939ac","#00b33c"),
      text.scale = 2, point.size
 = 4)
dev.off()
```


```{r}
path_table <- paste0("results/", name_ana, "/", 
                     name_ana, "_metabolites_randomforest.txt")

write.table(as.vector(unlist(list_cross_val)), file =path_table, sep = "\t", row.names = TRUE, 
            col.names = NA)
```


```{r}
path_table <- paste0("results/", name_ana, "/", 
                     name_ana, "_PLS-DA.txt")

write.table(as.vector(unlist(features_PLS_DA)), file =path_table, sep = "\t", row.names = TRUE, 
            col.names = NA)
```

```{r}
path_table <- paste0("results/", name_ana, "/", 
                     name_ana, "_metabolites_welshTtwotest.txt")

write.table(as.vector(unlist(results_man_withney$Accession)), file =path_table, sep = "\t", row.names = TRUE, 
            col.names = NA)
```

X16a.hydroxy.DHEA.3.sulfate
X2R.3R.dihydroxybutyrate
X4.hydroxyglutamate
X4.methylbenzenesulfonate
X5alpha.androstan.3alpha.17beta.diol.monosulfate..1.
X7.methylurate
andro.steroid.monosulfate.C19H28O6S..1..
androstenediol..3alpha..17alpha..monosulfate..3.
androsterone.sulfate
epiandrosterone.sulfate
hydroxypalmitoyl.sphingomyelin..d18.1.16.0.OH....
lamivudine
lysine
metabolonic.lactone.sulfate
methionine.sulfone
N1.Methyl.2.pyridone.5.carboxamide
N1.Methyl.4.pyridone.3.carboxamide
N1.methylinosine


### Overlap Cameroon/ India

Man Withney

1-palmitoyl-2-linoleoyl-GPC (16:0/18:2)
5alpha-androstan-3alpha,17beta-diol monosulfate (1)
5alpha-androstan-3beta,17beta-diol monosulfate (2)
5alpha-pregnan-3beta,20alpha-diol disulfate
7-methylurate
androstenediol (3alpha, 17alpha) monosulfate (3)
androsterone sulfate
bilirubin (E,Z or Z,E)*
efavirenz
epiandrosterone sulfate
glycohyocholate
lamivudine
metabolonic lactone sulfate
methionine sulfone


